---
title: "ADS_HW2"
author: "Ekaterina Hofrenning"
date: "3/28/2021"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_download: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```


```{r libraries}
# SEE modeldata package for new datasets
library(tidyverse)         # for graphing and data cleaning
library(tidymodels)        # for modeling
library(stacks)            # for stacking models
library(naniar)            # for examining missing values (NAs)
library(lubridate)         # for date manipulation
library(moderndive)        # for King County housing data
library(vip)               # for variable importance plots
library(DALEX)             # for model interpretation  
library(DALEXtra)          # for extension of DALEX
library(patchwork)         # for combining plots nicely
theme_set(theme_minimal()) # Lisa's favorite theme
```

```{r data}
data("lending_club")
# Data dictionary (as close as I could find): https://www.kaggle.com/wordsforthewise/lending-club/discussion/170691
```


**HW 2 GitHub link** : https://github.com/ehofrenning/datasci_hw2


\


## Modeling

*1. Explore the data, concentrating on examining distributions of variables and examining missing values.*

```{r}
data(lending_club)
lending_club
```


I first explore some main predictors via visualizations to examine the distributions:

```{r}
ggplot(lending_club, aes(x = funded_amnt)) +
  geom_density() +
  labs(title = "Funded Amount distribution") +
  xlab("Funded amount")

ggplot(lending_club, aes(x = int_rate)) +
  geom_histogram(bins = 40) +
  labs(title = "Interest rate distribution") +
  xlab("Interest rate")

ggplot(lending_club, aes(x = annual_inc)) +
  geom_histogram(bins = 150) +
  labs(title = "Annual income distribution") +
  xlab("Annual income")

# There are many more "good" loans than "bad"...:
ggplot(lending_club, aes(x = Class)) +
  geom_bar()
```


Here, I explore certain variables that I think may be all missing:

```{r}
lending_club %>%
  filter(acc_now_delinq == 1)

# THIS variable is all 0
lending_club %>%
  filter(delinq_amnt == 1)
```

From the above data exploration, I learned that I need to take out the *delinq_amnt* variable because all observations have a 0. 

*Do any data cleaning steps that need to happen before the model is build. For example, you might remove any variables that mean the same thing as the response variable (not sure if that happens here), get rid of rows where all variables have missing values, etc.*

```{r}
# Getting rid of delinq_amnt
lending <-
lending_club %>%
  select(-delinq_amnt)
```


```{r}
#Be sure to add more “bad” Classes. This is not the best solution, but it will work for now. (Should investigate how to appropriately use step_sample_up() function from themis).

# in an ideal world, we would want to upsample training data and apply it to untouched testing data... but this is fine.

create_more_bad <- lending %>% 
  filter(Class == "bad") %>% 
  sample_n(size = 3000, replace = TRUE)

lending_club_mod <- lending %>% 
  bind_rows(create_more_bad)
```


*3. Split the data into training and test, putting 75% in the training data.*

```{r}
set.seed(494) # for reproducibility

# split the data
lending_split <- initial_split(lending_club_mod, prop = .75, strata = Class)

# make training/test sets
lending_training <- training(lending_split)
lending_testing <- testing(lending_split)
```


*4. Set up the recipe and the pre-processing steps to build a lasso model. Some steps you should take:*
Make all integer variables numeric (I’d highly recommend using step_mutate_at() or this will be a lot of code). We’ll want to do this for the model interpretation we’ll do later.
Think about grouping factor variables with many levels.
Make categorical variables dummy variables (make sure NOT to do this to the outcome variable).
Normalize quantitative variables.

```{r}
# Make integers into numeric
lending_training[12:21] <- lapply(lending_training[12:21], as.numeric)
lending_training[9:10] <- lapply(lending_training[9:10], as.numeric)
lending_training[1] <- lapply(lending_training[1], as.numeric)

# Other pre-processing steps:
lending_recipe <- recipe(Class ~ ., data = lending_training) %>%
  step_normalize(all_numeric()) %>%   # normalize quant vars
  step_dummy(all_nominal(), -all_outcomes())   # make categorical vars into dummy EXCEPT for outcome

# check to see that this all worked
prep(lending_recipe) %>%
  juice() %>%
  head()
```


*5. Set up the lasso model and workflow. We will tune the penalty parameter.*

```{r}
# define lasso model
lending_lasso <- 
  logistic_reg(mixture = 1) %>%   # 1 indicates lasso
  set_engine("glmnet") %>%
  set_args(penalty = tune()) %>%   #we'll tune the lambda later
  set_mode("classification")
```

```{r}
# define LASSO workflow
lasso_wf <-
  workflow() %>%
  add_recipe(lending_recipe) %>%   #the processed data
  add_model(lending_lasso)     #the general lasso model
lasso_wf
```


*6. Set up the model tuning for the penalty parameter. Be sure to add the control_stack_grid() for the control argument so we can use these results later when we stack. Find the accuracy and area under the roc curve for the model with the best tuning parameter. Use 5-fold cv.*

```{r}
# 5-fold cv 
lending_cv <- vfold_cv(lending_training, v = 5)

# create penalty grid 
penalty_grid <- grid_regular(penalty(), levels = 10)

# create control grid (this we need for stacking i think)
ctrl_grid <- control_stack_grid()
```

```{r}
lending_lasso_tune <- 
  lasso_wf %>%
  tune_grid(resamples = lending_cv,
            grid = penalty_grid,
            control = ctrl_grid)

#choose best penalty
best_param <- lending_lasso_tune %>% 
  select_best(metric = "accuracy")

# finalize workflow
lending_lasso_final_wf <- lasso_wf %>% 
  finalize_workflow(best_param)

# fit final model with the best lambda
lending_lasso_final_mod <- lending_lasso_final_wf %>% 
  fit(data = lending_training)
```

```{r}
# average accuracy
lending_lasso_tune %>%
  collect_metrics() %>%
  filter(.metric == "accuracy") %>%
  summarise(mean_accuracy = mean(mean))
```

Now, I do the same thing but using the area under the ROC curve as the metric:
```{r}
# SAME THING BUT FOR ROC AUC

#best param
best_param2 <- lending_lasso_tune %>% 
  select_best(metric = "roc_auc")

# finalize workflow
lending_lasso_final_wf2 <- lasso_wf %>% 
  finalize_workflow(best_param2)

# fit final model with the best lambda
lending_lasso_final_mod2 <- lending_lasso_final_wf %>% 
  fit(data = lending_training)

# average roc_auc
lending_lasso_tune %>%
  collect_metrics() %>%
  filter(.metric == "roc_auc") %>%
  summarise(roc_auc = mean(mean))
```


```{r}
# predicting
# preds <- predict(lending_lasso_final_mod, lending_testing)
# preds

# add the predictions from the model onto the testing data
# test <-
lending_testing %>% 
  bind_cols(predict(lending_lasso_final_mod,
                    lending_testing, type = "prob")) %>%
    bind_cols(predict(lending_lasso_final_mod,
                    lending_testing)) 
test
```



*7. Set up the recipe and the pre-processing steps to build a random forest model. You shouldn’t have to do as many steps. The only step you should need to do is making all integers numeric*




*8. Set up the random forest model and workflow. We will tune the mtry and min_n parameters and set the number of trees, trees, to 100 (otherwise the next steps take too long).*



*9. Set up the model tuning for both the mtry and min_n parameters. Be sure to add the control_stack_grid() for the control argument so we can use these results later when we stack. Use only 3 levels in the grid. For the mtry parameter, you need to put finalize(mtry(), lending_training %>% select(-Class)) in as an argument instead of just mtry(), where lending_training is the name of your training data. This is because the mtry() grid will otherwise have unknowns in it. This part can take a while to run.*



*10. Find the best tuning parameters. What is the are the accuracy and area under the ROC curve for the model with those tuning parameters?*


*11. Use functions from the DALEX and DALEXtra libraries to create a histogram and boxplot of the residuals from the training data. How do they look? Any interesting behavior?*


*12. Use DALEX functions to create a variable importance plot from this model. What are the most important variables?*


*13. Write a function called cp_profile to make a CP profile. The function will take an explainer, a new observation, and a variable name as its arguments and create a CP profile for a quantitative predictor variable. You will need to use the predict_profile() function inside the function you create - put the variable name there so the plotting part is easier. You’ll also want to use aes_string() rather than aes() and quote the variables. Use the cp_profile() function to create one CP profile of your choosing. Be sure to choose a variable that is numeric, not integer. There seem to be issues with those that I’m looking into.*


*14. Use DALEX functions to create partial dependence plots (with the CP profiles in gray) for the 3-4 most important variables. If the important variables are categorical, you can instead make a CP profile for 3 observations in the dataset and discuss how you could go about constructing a partial dependence plot for a categorical variable (you don’t have to code it, but you can if you want an extra challenge). If it ever gives you an error that says, “Error: Can’t convert from VARIABLE to VARIABLE due to loss of precision”, then remove that variable from the list. I seem to have figured out why it’s doing that, but I don’t know how to fix it yet.*


*15. Fit one more model type of your choosing that will feed into the stacking model.*


*16. Create a model stack with the candidate models from the previous parts of the exercise and use the blend_predictions() function to find the coefficients of the stacked model. Create a plot examining the performance metrics for the different penalty parameters to assure you have captured the best one. If not, adjust the penalty. (HINT: use the autoplot() function). Which models are contributing most?*


*17. Fit the final stacked model using fit_members(). Apply the model to the test data and report the accuracy and area under the curve. Create a graph of the ROC and construct a confusion matrix. Comment on what you see. Save this final model using the saveRDS() function - see the Use the model section of the tidymodels intro. We are going to use the model in the next part. You’ll want to save it in the folder where you create your shiny app.*

